Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-07-29 draft.tex
added stuff on experiments chapter

We have to assess the model accuracy to extract key figures from it.
%%
The detection model depicts the vertebrae positions, but lacks detailed information about their volume.
%%
It captures reliable information about spine canal curvature.
%%
In a previous work \cite{KlemmVMV13, KlemmVIS14}, we extracted a centerline representation of the lumbar spine canal from the detection model (Fig.~\ref{fig:centerline} (a)).
%%
Using the Frenet formulas \cite{Frenet}, we extracted the following metrics from the model (Fig.~\ref{fig:centerline} (b)):
%%
%Thus, we extracted the following metrics from the model:
%%
%\paragraph{Curvature}
\begin{itemize}
	% \item The curvature of the centerline is calculated using the summed mean curvature between adjacent points:
% %%
% $\sum \frac{curvature_i}{curvature_all}$.
	\item \emph{Curvature} is calculated as weighted sum of curvature between all adjacent points describing the centerline: $\sum \frac{curvature_i}{curvature_all}$.
%%
	\item \emph{Torsion} (deviation of a curve from the course) is calculated as weighted sum of torsion between all adjacent points describing the centerline: $\sum \frac{torsion_i}{torsion_all}$.
%%
\end{itemize}
%%
These figures are also extracted in the sagital, coronal and transversal projection of the model.
%%
We assess the information gain of each dimension using these projections.
%%
This gives us a total of 8 image-derived parameters.

@@

\section{\uppercase{Experiments}}
\label{sec:Experiments}

\subsection{Preliminary Results}
\begin{itemize}
	\item Direkter Vergleich
	\item Verteilung des Means
	\item Korrelationsmatrix
	\item PCA
\end{itemize}
Weil das nicht ausreichend war, haben wir einen C5.0 Baum erstellt, um alle potentiellen Zusammenhänge zwischen bildbezogenen und Nichtbildbezogenen Parametern herzustellen - Was kann man mit den Bilddaten erklären und was nicht?

>>>>

We have to assess the model accuracy to extract key figures from it.
%%
The detection model depicts the vertebrae positions, but lacks detailed information about their volume.
%%
It captures reliable information about spine canal curvature.
%%
In a previous work \cite{KlemmVMV13, KlemmVIS14}, we extracted a centerline representation of the lumbar spine canal from the detection model (Fig.~\ref{fig:centerline} (a)).
%%
Using the Frenet formulas \cite{Frenet}, we extracted the following metrics from the model (Fig.~\ref{fig:centerline} (b)):
%%
%Thus, we extracted the following metrics from the model:
%%
%\paragraph{Curvature}
\begin{itemize}
	% \item The curvature of the centerline is calculated using the summed mean curvature between adjacent points:
% %%
% $\sum \frac{curvature_i}{curvature_all}$.
	\item \emph{Curvature} is calculated as weighted sum of curvature between all adjacent points describing the centerline: $\sum \frac{curvature_i}{curvature_all}$.
%%
	\item \emph{Torsion} (deviation of a curve from the course) is calculated as weighted sum of torsion between all adjacent points describing the centerline: $\sum \frac{torsion_i}{torsion_all}$.
	\item \emph{Curvature angle} is defined by the middle point of the spine canal centerline as \emph{vertex} and the line between middle point and top/bottom point as \emph{sides}.
%%
\end{itemize}
%%
These figures are also extracted in the sagittal, coronal and transversal projection of the model.
%%
We assess the information gain of each dimension using these projections.
%%
This gives us a total of 9 image-derived parameters.

@@

\section{\uppercase{Experiments}}
\label{sec:Experiments}
Spine shape is confounded by several somatometric variables.
%%
Larger people have also a longer spine and its shape is more straight compared to smaller subjects.
%%
Since men are on average taller than women, gender is another confounder.
%%
Large body weight increases the spine load, resulting in a bend shape.
%%
All these variables need to be taken into account, when correlating spine curvature and torsion with non-image parameter.
%%
Since the gender confounder mainly encodes body-height, we decided to divide subjects into body-size groups.
%%
To avoid small outlier groups, epidemiologists recommend using quartiles to discretize metric variables \cite{KlemmVIS14}.
\subsection{Preliminary Results}
\begin{figure}[!h]
  %\vspace{-0.2cm}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/image-parameter-range}
  \caption{image-parameter-range}
  \label{fig:image-parameter-range}
\end{figure}
%%
\noindent At first we assessed the image parameter.
%%
Figure~\ref{fig:image-parameter-range} (right) shows the range of each variable as box plot.
%%
The projections to the transversal planes attract attention as they have many outliers.
%%
We conclude that curvature is not as reliable on the transversal plane as it is on the other planes.
%%--------------------------------
\subsubsection{Correlation Matrix}
%%
We calculated an association matrix to to assess correlations between the image parameters.
%%
Curvature, curvature angle and torsion correlate with their planar projections.
%%
Also the mean curvature and the curvature angle correlate by a factor of -0.89.
%%
Torsion and curvature/curvature angle do not correlate.
%%--------------------------------------------------------
\subsubsection{Correlation of Image Parameters With Back Pain}
%%
As first experiment we correlated the shape parameter with the binary back pain indicator.
%%
No statistical significant correlation could be observed through all subject groups.
\com{Hier weitermachen. GGPLOT erklaeren.}

\begin{itemize}
	\item Direkter Vergleich
	\item Verteilung des Means
	\item Korrelationsmatrix
	\item PCA
\end{itemize}
Weil das nicht ausreichend war, haben wir einen C5.0 Baum erstellt, um alle potentiellen Zusammenhänge zwischen bildbezogenen und Nichtbildbezogenen Parametern herzustellen - Was kann man mit den Bilddaten erklären und was nicht?
